MCTS Based on Simple Regret

### 2012-06-05 paper/srcr.socs/paper.tex
added ucb root
socs final

\title{MCTS Based on Simple Regret}

@@

\maketitle

\begin{abstract}
UCT, a state-of-the art algorithm for Monte Carlo tree search (MCTS),
is based on UCB, a policy for the Multi-armed Bandit problem (MAB) that 
minimizes the cumulative regret.  However, search differs from MAB in
that in MCTS it is usually only the final ``arm pull''
that collects a reward, rather than all ``arm pulls''.
Therefore, it makes more sense to minimize the simple, rather than
cumulative, regret. We  introduce policies for
multi-armed bandits with lower simple
regret than UCB and develop a two-stage scheme (SR+CR) for MCTS
which outperforms UCT empirically. We also propose a sampling
scheme based on value of information (VOI), achieving an algorithm
that empirically outperforms other proposed algorithms.
\end{abstract}

@@

In adversarial search and optimization under uncertainty, the goal is
typically to either find a good strategy, or even just to
find the best first action of such a strategy, which is closer to the pure
exploration variant, as seen in the selection problem
\cite{Bubeck.pure,TolpinShimony.blinkered}. In the selection problem,
it is much better to minimize the \emph{simple} regret.  However, the
simple and the cumulative regret cannot be minimized
simultaneously \cite{Bubeck.pure}.

@@

@@

We examine some sampling schemes with super-polynomially
decreasing upper bounds on the simple regret. The bounds
suggest that these schemes achieve a lower simple regret
than uniform sampling; indeed, this is confirmed
by experiments. First, we consider $\varepsilon$-greedy sampling, which
pulls the arm that currently has the greatst sample mean, with probability
$0<\varepsilon<1$, and any other arm with probability $\frac {1-\varepsilon} {K-1}$. 
This sampling scheme exhibits an exponentially decreasing simple
regret. Moreover, as the number of arms $K$ grows, the bound for $\frac 1
2$-greedy sampling ($\varepsilon=\frac 1 2$) becomes considerably tighter than for uniform
random sampling ($\varepsilon=\frac 1 K$).

Another sampling scheme for simple regret minimization can be obtained
by distributing samples in a way similar to UCB, but sampling the current best arm
less often. This can be achieved by replacing $\log(\cdot)$ in
Equation~\ref{eqn:ucb} with $\sqrt\cdot$. This scheme also exhibits
a super-polynomially decreasing simple regret.

@@

However, for thee internal nodes optimizing simple regret is not the
answer, and cumulative regret optimization is not so far off the
mark. Our suggested improvement to UCT is the \textbf{SR+CR MCTS
sampling scheme} that combines different sampling schemes on the first
step and during the rest of each rollout.  We expect such two-stage
sampling schemes to outperform UCT.

@@

Further improvement can be achieved by computing or estimating the
VOI of the rollouts and choosing rollouts that
maximize the VOI. We estimate the VOI from
the current set of samples using distribution-independent concentration inequalities.
%\begin{eqnarray}
%VOI_\alpha&\approx& \frac {\overline X_\beta} {n_\alpha+1}
%\exp\left(-2(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\\
%VOI_i&\approx& \frac {1-\overline X_\alpha} {n_i+1}
%\exp\left(-2(\overline X_\alpha - \overline X_i)^2 n_i\right),\; i\ne\alpha\nonumber\\
%\mbox{where }&&\alpha=\arg\max_i \overline X_i,\quad
%             \beta=\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
%\end{eqnarray}
The ``VOI-aware'' scheme samples the action that has the maximum
estimated VOI.  Although this scheme appears too complicated for
a formal analysis, early experiments demonstrate a significantly lower
simple regret.

>>>>

\title{MCTS Based on Simple Regret$^*$}

@@

\maketitle
\insert\footins{\vspace{\baselineskip} \noindent\footnotesize Extended abstract of AAAI 2012 paper.}

@@

In adversarial search and optimization under uncertainty, the goal is
typically to either find a good strategy, or even just to
find the best first action of such a strategy. This is closer to the pure
exploration variant \cite{Bubeck.pure} as seen in the selection problem
\cite{TolpinShimony.blinkered}. In the selection problem,
it is much better to minimize the \emph{simple} regret.  However, the
simple and the cumulative regret cannot be minimized
simultaneously \cite{Bubeck.pure}.

@@

In the simple regret (selection) setting, the agent gets to collect
only the reward of the last pull. The \textbf{simple regret} $\IE r$
of a sampling policy for the Multi-armed Bandit Problem is the
expected difference between the best true expected reward $\mu_*$ and
the true expected reward $\mu_j$ of the arm with the greatest sample
mean, $j =\arg \max_i\overline X_i$:
\begin{equation}
\IE r=\sum_{j=1}^K\Delta_j\Pr(j=\arg\max_i\overline X_i)
\label{eqn:simple-regret}
\end{equation}
where $\Delta_j=\mu_*-\mu_j$.

@@

First, we consider $\varepsilon$-greedy sampling, which
pulls the arm that currently has the greatst sample mean, with probability
$0<\varepsilon<1$, and any other arm with probability $\frac {1-\varepsilon} {K-1}$. 
This sampling scheme exhibits an exponentially decreasing simple
regret. Moreover, as the number of arms $K$ grows, the bound for $\frac 1
2$-greedy sampling ($\varepsilon=\frac 1 2$) becomes considerably tighter than for uniform
random sampling ($\varepsilon=\frac 1 K$).

Another sampling scheme for simple regret minimization can be obtained
by distributing samples in a way similar to UCB, but sampling the current best arm
less often. This can be achieved by replacing $\log(\cdot)$ in
Equation~\ref{eqn:ucb} with $\sqrt\cdot$. This scheme (UCB$_{\sqrt{\cdot}}$) also exhibits
a super-polynomially decreasing simple regret.

@@

However, for thee internal nodes optimizing simple regret is not the
answer, and cumulative regret optimization is not so far off the
mark. Our suggested improvement to UCT is the \textbf{SR+CR MCTS
  sampling scheme}. The scheme selects an action at the current
root node according to a scheme suitable for minimizing the simple
regret (\textbf{SR}), such as $\frac 1 2$-greedy or
UCB$_{\sqrt{\cdot}}$, and then (at non-root nodes) selects actions
according to UCB, which approximately minimizes the cumulative regret
(\textbf{CR}).  We expect such two-stage sampling scheme to outperform
UCT.

@@

Further improvement can be achieved by computing or estimating the
value of information (VOI) of the rollouts and choosing rollouts that
maximize the VOI. We estimate the VOI from
the current set of samples using distribution-independent concentration inequalities.
%\begin{eqnarray}
%VOI_\alpha&\approx& \frac {\overline X_\beta} {n_\alpha+1}
%\exp\left(-2(\overline X_\alpha - \overline X_\beta)^2 n_\alpha\right)\\
%VOI_i&\approx& \frac {1-\overline X_\alpha} {n_i+1}
%\exp\left(-2(\overline X_\alpha - \overline X_i)^2 n_i\right),\; i\ne\alpha\nonumber\\
%\mbox{where }&&\alpha=\arg\max_i \overline X_i,\quad
%             \beta=\arg\max_{i,\,i\ne\alpha} \overline X_i\nonumber
%\end{eqnarray}
The ``VOI-aware'' scheme samples the action that has the maximum
estimated VOI.  Although this scheme appears too complicated for
a formal analysis, early experiments demonstrate a significantly lower
simple regret.
